---
title: "Companion file 1 for: "Towards a principled statistical workflow for the study of wildlife and environmental contaminants in the presence of method detection and quantification limits" "
format: html
---

This document is the first companion file for "Towards a principled statistical workflow for the study of wildlife and environmental contaminants in the presence of method detection and quantification limits". Here, we dive deeper in the concepts discussed in the manuscript.

```{r}
library(tidyverse)
library(brms)
library(cmdstanr)
library(tidybayes)
set.seed(333)
# Define and set the global theme
custom_theme <- theme(
  axis.line = element_line(linewidth = 2, lineend = "round"),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_text(size = 14, face = "bold"),
  axis.title = element_text(size = 16, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  legend.title = element_text(size = 14, face = "bold"),
  plot.title = element_text(size = 17, face = "bold"),
  plot.caption = element_text(face = "bold"),
  strip.background = element_rect(fill = "white"),
  strip.text = element_text(size = 12, face = "bold")
)

theme_set(custom_theme)
```

## Parameterizing the lognormal distribution
The lognormal distribution has two parameters, here defined as a and b 

Andrew MacDonald discusses moment matching for these two parameters in:
https://discourse.mc-stan.org/t/lognormal-regression-and-moment-matching/26557 
a
 is the median of the lognormal distribution. Its also the mean of the normal distribution you’d get if you took the log of these lognormal values. 
b
 is the standard deviation of those values.

Andrew provides functions to convert the mean and sd of a gaussian distribution to the a and b parameters

```{r}
calculate_a_lnorm <- function(mean, sd){

  log(mean^2/sqrt(sd^2 + mean^2))

}


calculate_b_lnorm <- function(mean, sd){

  sqrt(log(sd^2 / mean^2 + 1))

}
```

We then want to see what parameters brms models when fitting a hurdle lognormal model. We will include censored observations with a low probability of zeros to reduce the possible bias induced.


```{r}
# Set simulation parameters
n <- 2000          # Number of samples
mean_resp <- 0.23  # Average on reponse scale
sd_resp <- 0.06    # Standard devation on reponse scale
p_zeros <- 0.1     # Probability of zero
# Expected average including presence probability
expected_resp <- mean_resp * (1 - p_zeros) 
LOD <- 0.15        # Limit of detection
LOQ <- 0.19        # Limit of quantification
# Calculate a and b parameters of lognormal distribution
a <- calculate_a_lnorm(mean_resp, sd_resp)
b <- calculate_b_lnorm(mean_resp, sd_resp)
# Generate simulated data
tibble(
  # Generate lognormal concentrations
  conc = round(rlnorm(n, a, b),2),  
  # Presence probability of 0.9
  zeros = rbinom(n, prob = (1 - p_zeros), size = 1),  
  # Final concentrations (including zeros)
  Y = conc * zeros)  |>
  mutate(
    # format data
    y_cen = case_when(
    # Set values below LOD to LOD
      Y < LOD ~ LOD,  
    # Set values between LOD and LOQ to LOD
      Y >= LOD & Y < LOQ ~ LOD,  
    # Leave other values as is
      TRUE ~ Y),  
    # Create column for upper bound of interval censoring
    upper_int_cens = ifelse(Y >= LOD & Y < LOQ, LOQ, Y), 
    # Create column for censoring type
    censoring = case_when(
    # Left-censored
      Y < LOD ~ "left",  
    # Interval-censored
      Y >= LOD & Y < LOQ ~ "interval",  
    # Fully observed
      TRUE ~ "none")  
) -> censored_data   
```

We will fit a model to the data using the default {brms} priors for now just to see what parameters are modeled and on what scale


```{r, echo = TRUE}
bf_intercept <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1, 
                                                         hu ~ 1, 
                           family = hurdle_lognormal(link = "identity", 
                                                     link_sigma = "log", 
                                                     link_hu = "logit")
)

mod_intercept <- brms::brm(formula = bf_intercept,
                 data = censored_data,
                 iter = 2000, warmup = 1000, chains = 4, cores = 4,
                 seed = 333,
                 threads = threading(4, grainsize = 100),
                 backend = "cmdstanr")

tidybayes::epred_rvars(mod_intercept, 
            dpar = TRUE, 
            newdata = tibble(.rows = 1)) -> epred_intercept

epred_intercept
```

.epred seems similar to 'expected_resp'
mu = 'a' parameter
hu = 'p_zeros'
sigma = 'b' parameter

epred_rvars() thus does not give us the expected average of the observations without zeros. We can obtain that value from .epred/(1 - hu).
To get the sd of the observations without zeros, we can play around with the equation for parameter 'a' = log(mean^2/sqrt(sd^2 + mean^2)) to get
sd = mean * sqrt(mean^2 - exp(2a) / exp(2a))

```{r, echo = TRUE}
calculate_exp_mean <- function(epred, hu){

    epred / (1 - hu)

}

calculate_exp_sd<- function(mean, a){

    mean * sqrt((mean^2 - exp(2*a)) / exp(2*a))

}

# calculate quantities, pivot longer and visualize
epred_intercept |>
    mutate(exp_mean = calculate_exp_mean(epred = .epred, hu = hu),
           exp_sd = calculate_exp_sd(mean = exp_mean, a = mu)) |>
    pivot_longer(everything()) |>
    mutate(true_value = case_when(
        name == ".epred" ~ expected_resp,
        name == "mu" ~ a,
        name == "hu" ~ p_zeros,
        name == "sigma" ~ b,
        name == "exp_mean" ~ mean_resp,
        name == "exp_sd" ~ sd_resp
    ),
    name = case_when(
        name == ".epred" ~ "E[Y]",
        name == "mu" ~ "a",
        name == "hu" ~ "zero-probability",
        name == "sigma" ~ "b",
        name == "exp_mean" ~ "E[Y|presence]",
        name == "exp_sd" ~ "sd(Y)|presence"
    )) -> epred_intercept_long
    
    
ggplot(data = epred_intercept_long) +
    stat_slabinterval(aes(xdist = value, y = 250)) +
    geom_vline(aes(xintercept = true_value), 
               color = "red", 
               linewidth = 2, 
               lineend = "round") +
    facet_wrap(~name, scales = "free") +
    theme(axis.line.y = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_blank())
```

As we can see here, {brms} models the parameters a & b for the concentration. When using epred_ to get the expected parameter values, we get values for a and b that are hard to interpret but can be converted to the expected values of the mean and sd of the samples in which contaminants are present.

We must now think of priors for the a,b and hu parameters. As the a and b parameters are counter-intuitive, we can think of priors for the mean and sd of the observations and convert those priors to the a and b parameters. 

In this case, we know the true values of the parameters but in a real-world example, we would not. We can set up priors that are very wide for now. We may think that mean_resp could lie anywhere between 0 and 40, for example. A prior representing this belief could be Normal(20,10). We may also think that sd_resp is strictly positive and is about 10 on average. A prior representing this belief could be Exponential(0.1) where the mean of the exponential distribution is 1/0.1 = 10.


```{r}
tibble(
    prior_mean = rnorm(n, 20, 10),
    prior_sd = rexp(n,0.1),
    prior_a = calculate_a_lnorm(mean = prior_mean, sd = prior_sd),
    prior_b = calculate_b_lnorm(mean = prior_mean, sd = prior_sd)
) |>
    pivot_longer(everything()) |>
    mutate(true_value = case_when(
        name == "prior_mean" ~ mean_resp,
        name == "prior_sd" ~ sd_resp,
        name == "prior_a" ~ a,
        name == "prior_b" ~ b
    )) |>
    ggplot(aes(x = value)) +
    geom_density() +
    geom_vline(aes(xintercept = true_value), 
               color = "red", 
               linewidth = 2, 
               lineend = "round") +
    facet_wrap(~name, scales = "free")+
    theme(axis.line.y = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_blank())
    
```


To make our priors on a and b even more vague, we could choose a N(0, 2.5) for a and Exponential(1) for b.

For the hu parameter, we can set a wide prior of Normal(0,2) which on the probability scale is centered on 50% and goes near 0 and 100

Let's specify a model with these priors

```{r}
get_prior(bf_intercept, data = censored_data)

priors_intercept <- c(
    prior(normal(1, 2), class = "Intercept"),
    prior(normal(0, 1), class = "Intercept", dpar = "hu"),
    prior(exponential(3), class = "sigma")
)


prior_sim_intercept <- brms::brm(formula = bf_intercept,
                 data = censored_data,
                 prior = priors_intercept,
                 sample_prior = "only",
                 iter = 2000, warmup = 1000, chains = 4, cores = 4,
                 seed = 333,
                 threads = threading(4, grainsize = 100),
                 backend = "cmdstanr")

print(prior_sim_intercept)

# Visualize expectations
tidybayes::epred_rvars(prior_sim_intercept, 
            dpar = TRUE, 
            newdata = tibble(.rows = 1)) |>
    mutate(exp_mean = calculate_exp_mean(epred = .epred, hu = hu),
           exp_sd = calculate_exp_sd(mean = exp_mean, a = mu)) |>
    pivot_longer(everything()) |>
    mutate(true_value = case_when(
        name == ".epred" ~ expected_resp,
        name == "mu" ~ a,
        name == "hu" ~ p_zeros,
        name == "sigma" ~ b,
        name == "exp_mean" ~ mean_resp,
        name == "exp_sd" ~ sd_resp
    ),
    name = case_when(
        name == ".epred" ~ "E[Y]",
        name == "mu" ~ "a",
        name == "hu" ~ "zero-probability",
        name == "sigma" ~ "b",
        name == "exp_mean" ~ "E[Y|presence]",
        name == "exp_sd" ~ "sd(Y)|presence"
    )) |>
    ggplot() +
    stat_slabinterval(aes(xdist = value, y = 250)) +
    geom_vline(aes(xintercept = true_value), 
               color = "red", 
               linewidth = 2, 
               lineend = "round") +
    facet_wrap(~name, scales = "free") +
    theme(axis.line.y = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_blank())
```

After playing around with some values, we can choose these priors as they appear wide enough to be vaguely informative. This may not be an optimal approach however. It is also pertinent to simulate observations from this model and compare them to our data.

