---
title: "Companion file 2 for: "Towards a principled statistical workflow for the study of wildlife and environmental contaminants in the presence of method detection and quantification limits" "
format: html
---
In this document, we code custom prior/posterior predictive checks for hurdle left- and interval-censored models. The goal is to bin the predicted number of observations into zeros, left censored, interval censored, quantified categories. For predictions above LOQ (quantified concentrations), we will also plot the probability density and cumulative probability density functions. In all cases, we will compare the predicted observations to the observed ones to assess model fit.

First, we will simulate some hurdle left and interval censored lognormal data.

```{r}
library(tidyverse)
library(brms)
library(cmdstanr)
library(tidybayes)
library(patchwork)
library(furrr)
library(httpgd)
set.seed(333)
# Define and set the global theme
custom_theme <- theme(
  axis.line = element_line(linewidth = 2, lineend = "round"),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_text(size = 14, face = "bold"),
  axis.title = element_text(size = 16, face = "bold"),
  legend.text = element_text(size = 12, face = "bold"),
  legend.title = element_text(size = 14, face = "bold"),
  plot.title = element_text(size = 17, face = "bold"),
  plot.caption = element_text(face = "bold"),
  strip.background = element_rect(fill = "white"),
  strip.text = element_text(size = 12, face = "bold")
)

theme_set(custom_theme)

calculate_a_lnorm <- function(mean, sd){

  log(mean^2/sqrt(sd^2 + mean^2))

}


calculate_b_lnorm <- function(mean, sd){

  sqrt(log(sd^2 / mean^2 + 1))

}

hgd()
```



```{r}
# Set simulation parameters
n <- 2000          # Number of samples
mean_resp <- 0.23  # Average on reponse scale
sd_resp <- 0.06    # Standard devation on reponse scale
p_zeros <- 0.1     # Probability of zero
# Expected average including presence probability
expected_resp <- mean_resp * (1 - p_zeros) 
LOD <- 0.15        # Limit of detection
LOQ <- 0.19        # Limit of quantification
# Calculate a and b parameters of lognormal distribution
a <- calculate_a_lnorm(mean_resp, sd_resp)
b <- calculate_b_lnorm(mean_resp, sd_resp)
# Generate simulated data
tibble(
  # Generate lognormal concentrations
  conc = round(rlnorm(n, a, b),2),  
  # Presence probability of 0.9
  zeros = rbinom(n, prob = (1 - p_zeros), size = 1),  
  # Final concentrations (including zeros)
  Y = conc * zeros,
  contaminant = "Fake")  |>
  mutate(
    # format data
    y_cen = case_when(
    # Set values below LOD to LOD
      Y < LOD ~ LOD,  
    # Set values between LOD and LOQ to LOD
      Y >= LOD & Y < LOQ ~ LOD,  
    # Leave other values as is
      TRUE ~ Y),  
    # Create column for upper bound of interval censoring
    upper_int_cens = ifelse(Y >= LOD & Y < LOQ, LOQ, Y), 
    # Create column for censoring type
    censoring = case_when(
    # Left-censored
      Y < LOD ~ "left",  
    # Interval-censored
      Y >= LOD & Y < LOQ ~ "interval",  
    # Fully observed
      TRUE ~ "none")  
) -> censored_data
```

Model
We will use brms default priors for the purpose of this document
```{r}
bf_intercept <- bf(y_cen | cens(censoring , upper_int_cens) ~ 1, 
                                                         hu ~ 1, 
                           family = hurdle_lognormal(link = "identity", 
                                                     link_sigma = "log", 
                                                     link_hu = "logit")
)

mod_intercept <- brms::brm(formula = bf_intercept,
                 data = censored_data,
                 iter = 500, warmup = 100, chains = 1, # we usually want more tierations, warmups and 4 chains but this will speed up computation for this example
                 seed = 333,
                 backend = "cmdstanr")
```

We now want to extract predicted values from this model to compare with the observed values. 


```{r}
censored_data |>
    # add predicted obervations
    tidybayes::add_predicted_draws(mod_intercept, ndraws = 100) |> 
    # creates a grouped tibble so we need to ungroup
    ungroup() |>
    # now we bin values into < LOD, > LOD <= LOQ, and quantified
    mutate(
    pred_type = case_when(
    # Left-censored
      .prediction < LOD ~ "left",  
    # Interval-censored
      .prediction  >= LOD & .prediction  < LOQ ~ "interval",    
    # Fully observed
      TRUE ~ "none" 
    ))  -> predicted_values
```

Now that we have predicted values in their respective bins, we can start comparing them to the actual observations.

First, we count the number of observations in each bin

```{r}
predicted_values |>
    # Each actual observation is gets 100 draws from the posterior 
    # so we summarize by draw
    group_by(.draw) |>
    # Count the number of predicted obervations per bin
    count(pred_type) |>
    ungroup() -> binned_predictions

# Do the same for the actual data

censored_data |>
    count(censoring) -> binned_observations
```

Plot bins

```{r}
ggplot() + 
    geom_col(data = binned_observations, 
            aes(x = factor(censoring, levels = c("left", "interval", "none")), 
                y = n), 
            fill  = "darkblue") + 
    stat_pointinterval(data = binned_predictions, 
                       aes(x = factor(pred_type, levels = c("left", "interval", "none")), 
                           y = n), 
                       color = "lightblue", size = 1) + 
    labs(x = NULL, y = "n observations")
```

We can see that the model puts a bit too many observations in the interval bin but the actual number of observations in this bin is within the 95% interval 

While this is informative, we also want to compare the distribution of quantified values.


```{r}
ggplot() +
    geom_density(data = predicted_values |> filter(pred_type == "none"), 
                 aes(x = .prediction, group = .draw), 
                 color = "lightblue", 
                 trim = TRUE) +
    geom_density(data = censored_data |> filter(censoring == "none"), 
                 aes(x = y_cen), 
                 color = "darkblue", 
                 linewidth = 1, 
                 trim = TRUE) +
    labs(x = "quantified values")
```

Even though the model recovers the parameters well, the posterior predictive check is not optimal.

We can also use the ecdf as another version of the ppc


```{r}
ggplot() +
    stat_ecdf(data = predicted_values |> filter(pred_type == "none"),
              aes(x = .prediction, group = .draw), 
              color = "lightblue") + 
    stat_ecdf(data  = censored_data |> filter(censoring == "none"), 
              aes(x = y_cen), 
              color = "darkblue", 
              linewidth = 1) +
    labs(x = "quantified values")
```

We might also be interested in the tails of the distribution. So we will compare the 95th percentile.


```{r}
ggplot() +
    geom_vline(data = censored_data |> 
                      filter(censoring == "none") |>
                      summarise(perc_95 = quantile(y_cen, 0.95)),
    aes(xintercept = perc_95), 
    linewidth = 1, 
    color = "darkblue") +
    stat_pointinterval(data = predicted_values |> 
                              filter(pred_type == "none") |>
                              group_by(.draw) |>
                              summarise(perc_95 = quantile(y_cen, 0.95)) |>
                              ungroup(), 
                       aes(x = perc_95), 
                       color = "lightblue")
```

It seems like the predicted 95th is capped at 0.34 which may be due to the default brms priors. We will see how these diagnostics change in Companion File 1 when playing around with different priors. 

We have explored here three diagnostic methods to assess model fit. We can now convert these into functions that will also apply to models with multiple contaminants.

Load model
THIS WILL CHANGE

```{r}
test_long_data <- read_rds(file = "test_long_data.Rda")

dl <- test_long_data %>%
  mutate(lat = as.numeric(lat), lon = as.numeric(lon)) %>%
  rowwise() %>%
  mutate(tot_wat_1 = Water.1000 + Wetland.1000) %>%
  ungroup() %>%
  mutate(scl_water = scale(tot_wat_1),
         year = as.factor(year),
         farm_ID = as.factor(farm_ID),
         pesticide = as.factor(pesticide))

dl_sub <- dl %>%
  mutate(pesticide = as.character(pesticide)) %>%
  filter(!pesticide %in% c("bentazon", "carbaryl", "parathion", "piperonyl.butoxide"))

dl_sub2 <- subset(dl_sub, pesticide != "one.naphthol")

dl_all <- dl %>%
  mutate(pesticide = as.character(pesticide)) %>%
  subset( pesticide != "one.naphthol") %>%
  mutate(pesticide = as.factor(pesticide))

mod_pest_agri_wat_mu_hu_all <- brm(data = dl_all,
           formula = bf_pest_agri_wat_mu_hu ,
           prior = prior_pest_agri_wat_mu_hu_all ,
           iter = 15000, warmup = 2000, chains = 4, cores = 4,
           threads = threading(4, grainsize = 100),
           backend = "cmdstanr",
           control = list(adapt_delta = 0.99),
           file = "mod_pest_agri_wat_mu_hu_all")
```


Function to generate predictions and bin them by censoring category
```{r}
ppc_predict <- function(data,            # Data used in model
                        model,           # model
                        draws = 100,     # number of posterior draws
                        re_form          # Hierarchical effects included
                        ) {

data |>
    # add predicted obervations
    tidybayes::add_predicted_draws(model, ndraws = draws, re_formula = re_form) |> 
    # creates a grouped tibble so we need to ungroup
    ungroup() |>
    # now we bin values into < LOD, > LOD <= LOQ, and quantified
    mutate(
    pred_type = factor(
      case_when(
    # Left-censored
      .prediction < LOD ~ "left",  
    # Interval-censored
      .prediction  >= LOD & .prediction  < LOQ ~ "interval",    
    # Fully observed
      TRUE ~ "none"),
      levels = c("left", "interval", "none")) )
}

# Test with multiple contaminants
test_pred <- ppc_predict(data = dl_all, 
                         model = mod_pest_agri_wat_mu_hu_all, 
                         draws = 2,
                         re_form = ~ (0 +  pesticide | farm_ID) + (1|year) + (1|pesticide))

# Test with single contaminant
pred_single <- ppc_predict(data = censored_data, 
                           model = mod_intercept, 
                           draws = 2,
                           re_form = NA)
```


Density plot with binned censoring type

```{r}
cens_dens <- function(           data,            # Data used in model
                                 preds,           # Model predictions generated by ppc_predict()
                                 cont_col,        # Column for contaminant ID
                                 cens_col,        # Column specifying censoring type in data
                                 y,               # Column specifying (censored) observations in data
                                 limits = FALSE   # Crop plot to the limit of observed data?
                           ) {
  # Bin predictions     
  binned_preds <- preds |>
      group_by(!!sym(cont_col), .draw) |>
      count(pred_type) |>
      ungroup()

  # Bin observations
  binned_obs <- data |>
      mutate(!!sym(cens_col) := factor(!!sym(cens_col), levels = c("left", "interval", "none"))) |>
      group_by(!!sym(cont_col)) |>
      count(!!sym(cens_col)) |>
      ungroup()
 
  # Get unique contaminants
  contaminants <- unique(data %>% pull(!!sym(cont_col)))

  # Set up parallel processing
  plan(multisession)
  
  # Create and collect plots for each contaminant in parallel
  plot_list <- future_map(contaminants, function(cont) {
      # Filter data for current contaminant
      binned_obs_cont <- binned_obs |> filter(!!sym(cont_col) == cont)
      binned_preds_cont <- binned_preds |> filter(!!sym(cont_col) == cont)
      data_cont <- data |> filter(!!sym(cont_col) == cont)
      pred_cont <- preds |> filter(!!sym(cont_col) == cont)
      
      # Create categorical plot
      plot_cat <- ggplot() + 
        geom_col(data = binned_obs_cont, 
                   aes(x = !!sym(cens_col), y = n), 
                   fill = "darkblue") + 
        stat_pointinterval(data = binned_preds_cont, 
                             aes(x = pred_type, y = n), 
                             color = "lightblue", size = 1) + 
        theme_classic() +
        labs(x = NULL, y = "n observations")

      # Create quantitative plot
      plot_quant <- ggplot() +
        geom_density(data = pred_cont |> filter(pred_type == "none"), 
                          aes(x = .prediction, group = .draw), color = "lightblue", 
                          trim = TRUE) +
        geom_density(data = data_cont |> filter(!!sym(cens_col) == "none"), 
                          aes(x = !!sym(y)), color = "darkblue", linewidth = 1, 
                          trim = TRUE) +
        theme_classic() + 
        labs(x = "quantified values")

      # Add limits if specified
      if(limits == TRUE) {
        plot_quant <- plot_quant + 
            coord_cartesian(xlim = c(0, max(data_cont |> 
                                            pull(!!sym(y)), 
                                            na.rm = TRUE)))
      }
      
      # Combine plots and add title
      combined_plot <- plot_cat + plot_quant + 
        plot_layout(widths = c(1, 3)) + 
        plot_annotation(title = paste(cont))
      
      return(combined_plot)
  })
  
  # Clean up parallel processing
  plan(sequential)
  
  # Print plots sequentially
  walk(plot_list, print)
}

# Test with multiple contaminants
cens_dens(data = dl_all, preds = test_pred, cont_col = "pesticide", cens_col = "censoring", y = "y_cen")

# Test with single contaminant
cens_dens(data = censored_data, preds = pred_single, cont_col = "contaminant", cens_col = "censoring", y = "y_cen")
```

Function with ECDF

```{r}
cens_ecdf <- function(           data,            # Data used in model
                                 preds,           # Model predictions generated by ppc_predict()
                                 cont_col,        # Column for contaminant ID
                                 cens_col,        # Column specifying censoring type in data
                                 y,               # Column specifying (censored) observations in data
                                 limits = FALSE   # Crop plot to the limit of observed data?
                           ) {
  # Bin predictions     
  binned_preds <- preds |>
      group_by(!!sym(cont_col), .draw) |>
      count(pred_type) |>
      ungroup()

  # Bin observations
  binned_obs <- data |>
      mutate(!!sym(cens_col) := factor(!!sym(cens_col), levels = c("left", "interval", "none"))) |>
      group_by(!!sym(cont_col)) |>
      count(!!sym(cens_col)) |>
      ungroup()
 
  # Get unique contaminants
  contaminants <- unique(data %>% pull(!!sym(cont_col)))

  # Set up parallel processing
  plan(multisession)
  
  # Create and collect plots for each contaminant in parallel
  plot_list <- future_map(contaminants, function(cont) {
      # Filter data for current contaminant
      binned_obs_cont <- binned_obs |> filter(!!sym(cont_col) == cont)
      binned_preds_cont <- binned_preds |> filter(!!sym(cont_col) == cont)
      data_cont <- data |> filter(!!sym(cont_col) == cont)
      pred_cont <- preds |> filter(!!sym(cont_col) == cont)
      
      # Create categorical plot
      plot_cat <- ggplot() + 
        geom_col(data = binned_obs_cont, 
                   aes(x = !!sym(cens_col), y = n), 
                   fill = "darkblue") + 
        stat_pointinterval(data = binned_preds_cont, 
                             aes(x = pred_type, y = n), 
                             color = "lightblue", size = 1) + 
        theme_classic() +
        labs(x = NULL, y = "n observations")

      # Create quantitative plot
      plot_quant <- ggplot() +
        stat_ecdf(data = pred_cont |> filter(pred_type == "none"), 
                          aes(x = .prediction, group = .draw), color = "lightblue") +
        stat_ecdf(data = data_cont |> filter(!!sym(cens_col) == "none"), 
                          aes(x = !!sym(y)), color = "darkblue", linewidth = 1) +
        theme_classic() + 
        labs(x = "quantified values")

      # Add limits if specified
      if(limits == TRUE) {
        plot_quant <- plot_quant + 
            coord_cartesian(xlim = c(0, max(data_cont |> 
                                            pull(!!sym(y)), 
                                            na.rm = TRUE)))
      }
      
      # Combine plots and add title
      combined_plot <- plot_cat + plot_quant + 
        plot_layout(widths = c(1, 3)) + 
        plot_annotation(title = paste(cont))
      
      return(combined_plot)
  })
  
  # Clean up parallel processing
  plan(sequential)
  
  # Print plots sequentially
  walk(plot_list, print)
}

# Test with multiple contaminants
cens_ecdf(data = dl_all, preds = test_pred, cont_col = "pesticide", cens_col = "censoring", y = "y_cen")

# Test with single contaminant
cens_ecdf(data = censored_data, preds = pred_single, cont_col = "contaminant", cens_col = "censoring", y = "y_cen")
```

95th percentile

```{r}
cens_95th <- function(           data,            # Data used in model
                                 preds,           # Model predictions generated by ppc_predict()
                                 cont_col,        # Column for contaminant ID
                                 cens_col,        # Column specifying censoring type in data
                                 y                # Column specifying (censored) observations in 
                           ) {

 
  # Get unique contaminants
  contaminants <- unique(data %>% pull(!!sym(cont_col)))

  # Set up parallel processing
  plan(multisession)
  
  # Create and collect plots for each contaminant in parallel
  plot_list <- future_map(contaminants, function(cont) {
      # Filter data for current contaminant
      data_cont <- data |> filter(!!sym(cont_col) == cont)
      pred_cont <- preds |> filter(!!sym(cont_col) == cont)

      # Plot
      plot <- ggplot() +
              geom_vline(data = data_cont |> 
                                filter(!!sym(cens_col) == "none") |>
                                summarise(perc_95 = quantile(!!sym(y), 0.95, na.rm = TRUE)),
                         aes(xintercept = perc_95), 
                         linewidth = 1, 
                         color = "darkblue") +
              stat_pointinterval(data = pred_cont |> 
                                        filter(pred_type == "none") |>
                                        group_by(.draw) |>
                                        summarise(perc_95 = quantile(.prediction, 0.95, na.rm = TRUE)) |>
                                        ungroup(), 
                                 aes(x = perc_95), 
                                 color = "lightblue") +
              theme_classic() +
              labs(x = "95th percentile",
                   title = paste(cont)) +
              theme(axis.line.y = element_blank(),
                    axis.text.y = element_blank(),
                    axis.title.y = element_blank(),
                    axis.ticks.y = element_blank())
      
      return(plot)
  })
  
  # Clean up parallel processing
  plan(sequential)
  
  # Print plots sequentially
  walk(plot_list, print)
}

# Test with multiple contaminants
cens_95th(data = dl_all, preds = test_pred, cont_col = "pesticide", cens_col = "censoring", y = "y_cen")

# Test with single contaminant
cens_95th(data = censored_data, preds = pred_single, cont_col = "contaminant", cens_col = "censoring", y = "y_cen")
```